### 2022-06-22 src/ms.tex
Finished Chapter 2 draft

\subsection{What about machine learning?}
%% mention that even if we have a model focused on inference, it might be best if we judge 
%% its quality through its predictive performance
%
\section{Automatic differentiation}
%\section{Putting it all together -- \emph{probabilistic programming}}
%\label{sec:programming}
%\subsection{The success of machine learning}
%% that quote from Yann LeCun on differentiable code 
%\subsection{Automatic differentiation}
%\subsubsection{Forward mode autodiff}
%\subsubsection{Reverse mode autodiff or backpropagation}
%\subsubsection{Jacobian vector products}
%\subsection{JAX}
%\subsection{Numpyro}

% CHAPTER 3: Modeling microlensing events 
%% PLOT POINTWISE LOO 
\chapter{Modeling microlensing events}
\label{ch:microlensing}
\section{Single lens events}
\section{Modeling single lens events}
\section{Numerical solutions to the lens equation}
\section{\textsf{caustics} -- computing the magnification of an extended limb-darkened source}
\section{Comparison with previous work}
\section{Modeling binary lens events}
\section{Population-level inference using hierarchical modeling}

% CHAPTER 4: Mapping the surface of Io 
\chapter{Mapping the surface of Io}
\label{ch:mapping_io}
\section{The data}
\section{Information content}
\section{Pixel sampling}
\section{A static map model}
\section{A dynamic map model}

% CHAPTER 5: Mapping the surfaces of exoplanets
\chapter{Mapping the surfaces of exoplanets}
\label{ch:mapping_exoplanets}
\section{Previous work}
\section{What can we learn using eclipse mapping?}
\section{Time dependent maps}

% CHAPTER 6: Conclusion
\chapter{Conclusion}
\section{Summary of the thesis}
\subsection{Microlensing}
\subsection{Occultation mapping}
\section{Future work}
\subsection{Microlensing in the era of the Roman telescope}
\subsection{Mapping volcanic activity on Io}
\subsection{Mapping exoplanets with JWST and beyond}

>>>>

\subsection{What about machine learning?}
How is \textsl{machine learning} (ML) different from the topics we have discussed so far? 
I would say the key difference is that ML and particularly \textsl{deep learning} (DL) 
\emph{ML is much more focused on predicting future data} conditional on observing 
past data  than on estimating the parameters of some model, which is the subject of 
statistical inference. Classic example is that of neural networks which 
today routinely consist of millions, even billions of parameters (weights) but 
the  values of these weights are not important, only the prediction error 
(on unseen data) is what matters. For this reason optimization methods such as 
\textsl{stochastic gradient descent} (SGD) which are computationally much 
cheaper to implement are dominant over sampling methods such as MCMC.

Despite the fact that the deep learning revolution started in 2012, 
deep learning has not had as much impact  in the physical sciences as many have 
expected  despite the fact that  there are many papers on the topic.
DL methods tend to be black boxes which are very efficient at 
the task they were trained to do, but their internal structure is usually 
illegible. This is not as much of an issue in industry applications as it is 
in the sciences where we care deeply about the precise values of very few 
parameters.
Nevertheless, all of this is changing rapidly and there is  a lot of work 
going on the boundary of physics and machine learning. One example is the field 
of \textsl{likelihood free inference} (LFI) which aims to emulate expensive 
physics simulators (such as massive hydrodynamics simulations) with neural 
networks so that one can do inference with much faster emulators 
\citep[see for example the review paper by]{2020PNAS..11730055C}. There has 
also been a lot of work on neural networks which respect certain physical 
symmetries (see for example Hamiltonian and Lagrangian neural 
networks\footnote{Or \textsl{geometric deep learning} more generally.})
and even automated discovery of succinct mathematical descriptions of 
the dynamical behavior of a physical system through observations of its 
behaviour \citep{arXiv:2006.11287}.

However, the biggest significance of the atmospheric rise of machine learning 
in my opinion is that the computational infrastructure developed for training 
large neural networks can also be used for scientific computing. In particular,
two of the key technologies responsible for the deep learning  revolution
are \textsl{Graphical Processing Units} (GPUs) and 
\textsl{automatic differentiation} (AD), are also very useful 
for building scientific models.

\section{Automatic differentiation}
\label{sec:autodiff}
\textsl{Automatic differentiation} (AD or autodiff for short)\citep{10.1145/355586.364791} is an old idea 
that has revolutionized the fields of machine learning, AI and optimization in the 
recent decade. 
The key idea behind automatic differentiation 
is that any model implemented in computer code (such as \textsf{C++} or 
\textsf{Python}) is fundamentally a composition 
of simple functions such as additions, multiplications and trig functions 
which can be easily differentiated and these gradients evaluated at each step 
of the computation can be combined using the chain rule from calculus to 
yield the exact (up to machine precision) derivative of the complete model.
Specialized automatic differentiation libraries build a representation of the 
code as graph, evaluate the derivatives at each node and propagate them through 
the entire graph.

A variant of automatic differentiation called reverse mode automatic differentiation 
is widely used in machine learning (where it goes under the name of 
\emph{backpropagation}) to differentiate scalar functions with respect to 
millions if not billions of parameters. 
Backpropagation and autodiff in general tells us how to change the inputs of a 
model (even if there are billions of inputs) to obtain a particular output.
Autodiff has been used to differentiate through all sorts of things, including 
complex simulations of physical systems such as molecular dynamics simulations,
\citep{arXiv:1912.04232}, cosmological simulations \citep{2016MNRAS.463.2273F},
 ODE and PDE solvers etc. It has even been applied to differentiating code at 
 the level of a compiler using the LLVM compiler infrastructure \citep{arXiv:2010.01709}.
 Because the wide variatey of autodiff applications
the famous AI researcher Yann LeCun recently popularized the term 
\textsl{differentiable programming} to refer to any computer code which is 
automatically differentiable.

The reason why I am discussing this topic in the first place is that implementing AD 
for astrophysical models enables the use  of advanced MCMC methods such as 
Hamiltonian Monte Carlo and much faster and more stable numerical optimization  
(for obtaining the MLE estimate for example).

\subsection{Three ways of differentiating computer programs}
There are three ways of computing derivatives:
\begin{itemize}
    \item \textbf{Symbolic differentiation}: computing \emph{exact} derivatives of 
    functions using either a pen and paper or a computer algebra system such as 
    \textsf{Wolfram Mathematica} or \textsf{SymPy} \citep{10.7717/peerj-cs.103}.
    \item \textbf{Numerical differentiation}: computing \emph{approximate} derivatives 
    using \textsl{finite difference}  methods.
    \item \textbf{Automatic differentiation}: computing \emph{exact} derivatives of entire 
    computer programs containing things like conditionals, loops, etc.
\end{itemize}
Symbolic differentiation is familiar to everyone from basic calculus. Using either 
pen and paper  or Computer Algebra Systems (CAS) we can compute exact 
derivatives of functions. 
If our model is expressible as a straightforward closed-form function it is 
often not to difficult to compute the derivatives of interest and indeed this 
is what people have been doing for decades to improve the performance of their
models\footnote{Before the advent of easy to use open-source AD libraries 
the contribution of many papers in applied AI would be to compute the analytic 
derivatives for some specific model.}

The challenge with symbolic differentiation is that it is not always possible 
to easily compute the gradients, especially if the model is implemented in terms 
of a code which contains control flow constructs such as loops and conditionals, 
or iterative solvers. In those cases gradients of the output of the model 
are with respect to the input are usually computed using finite-difference 
methods. In finite-difference methods the derivative of some function $f(x)$
is defined as 
\begin{equation}
    f^\prime(x)=\lim _{h \rightarrow 0} \frac{f(x+h / 2)-f(x-h / 2)}{h}
\end{equation}
for some small step size $h$. Numerical gradients are easy to compute even with 
complicated models but they have two key drawbacks. First, obtaining a 
good approximation of the derivatives requires a small step size $h$ which is 
can lead to errors due to numerical cancellations.
The second is that if we want to evaluate the Jacobian of a function with lots 
of input parameters we would need to compute a finite difference gradient 
for each partial derivative which is prohibitively expensive in high dimensions.

Automatic differentiation solves these problems. If AD is properly implemented 
evaluating the exact gradient is only a factor of a few slower than evaluating 
the function itself, even for models with millions of parameters.
The drawback is that it is usually
necessary to express our model in terms of constructs which are implemented in 
specialized AD libraries. It also requires a slightly different programming style
than what most people are used to. Fortunately there are now lots very flexible 
and easy to use high performance AD libraries.

\subsection{Forward and reverse-mode autodiff and the chain rule}
What follows is mostly a summary of the chapter on autodiff 
from \citet{murphy_book_2023}. First, we define the notation
\begin{equation}
\partial x_1 f\left(x_{1}, x_{2}, \ldots, x_{n}\right)
\end{equation} 
which denotes the partial derivative of function 
$f\,:\,\mathbb{R}^{n} \rightarrow \mathbb{R}$ with respect to argument $x_1$ 
evaluated at the point $(x_{1}, x_{2}, \ldots, x_{n})$.
We define the derivative operator $\partial$ and denote the derivative of a 
differentiable function $f\,:\,\mathbb{R}^{n} \rightarrow \mathbb{R}^m$ as 
\begin{equation}
    \partial f\,:\,\mathbb{R}^{n} \rightarrow \mathbb{R}^{m\times n}
    \label{eq:general_derivative}
\end{equation}
The function $\partial f$ is a \textsl{linear map} which maps a point 
$\vec x\in \mathbb{R}^n$ to the \textsl{Jacobian} of all partial derivatives 
evaluated at $\vec x$.  $\partial f$ evaluated at point $\vec x$,
$\partial f (\vec x)$ is the \textsl{linearization} of function $f$ at point $\vec x$.
The Jacobian can be be represent as a matrix $\vec J$ whose 
elements $J_{ij}$ are given by 
\begin{equation}
J_{i j}=\frac{\partial f_{i}}{\partial x_{j}}
\end{equation}
  To make this slighly less abstract, consider the case when
$m=n=1$, the function takes a scalar and returns a scalar, then 
Equation~\ref{eq:general_derivative} is just the familiar derivative $f'(x)$:
\begin{equation}
    f^\prime\,:\,\mathbb{R} \rightarrow \mathbb{R}
\end{equation}
If the $f$ takes a vector and returns a scalar ($m=1$) then derivative operator is 
the \textsl{gradient} $\nabla f$:
\begin{equation}
    \partial f\,:\,\mathbb{R}^{n} \rightarrow \mathbb{R}^{1\times n}
\end{equation}

We now introduce two key ingredients for an automatic differentiation framework 
which enable us to efficiently compute the Jacobian for functions of arbitrary 
complexity.

\subsubsection{Jacobian-vector products (JVPs)}
Now consider consider a perturbation $\vec v\in\mathbb{R}^n$ to the function input 
$\vec x$. The  mapping 
\begin{equation}
(\vec{x}, \vec{v}) \mapsto \partial f(\vec{x})\,\vec v
\label{eq:jvp}
\end{equation}
is called the \textsl{Jacobian-vector product} (JVP). For example, if $f=\cos(x)$ 
for a scalar $x$ and $v$ is also a scalar, the JVP would be 
\begin{equation}
(x, v) \mapsto \partial \cos (x) v=-v \sin (x)\hquad .
\end{equation}
Why is this useful? With a function $f\,:\,\mathbb{R}^n\rightarrow\mathbb{R}^m$
and a \textsl{one-hot} vector $\vec v=(1, 0, \ldots, 0)^\intercal$ we can use  the JVP to 
evaluate the Jacobian matrix one column at a time:
\begin{equation}
    \partial f(\vec x)\,\vec v = 
    \begin{pmatrix}
\frac{\partial f_{1}}{\partial x_{1}} & \cdots & \frac{\partial f_{1}}{\partial x_{n}} \\
\vdots & \ddots & \vdots \\
\frac{\partial f_{m}}{\partial x_{1}} & \cdots & \frac{\partial f_{m}}{\partial x_{n}}
    \end{pmatrix}
    \begin{pmatrix}
        1\\
        0\\
        \vdots\\
        0
    \end{pmatrix}
    = \begin{pmatrix}
        \frac{\partial f_{1}}{\partial x_{1}}\\
        \frac{\partial f_{2}}{\partial x_{1}}\\
        \vdots\\
        \frac{\partial f_{m}}{\partial x_{1}}
    \end{pmatrix}
\end{equation}
Evaluating one column of the Jacobian in this way costs about the same as one function evaluation
so this is efficient for fucntions with ``tall Jacobians'' (functions with many outputs and few
inputs) but innefficient for functions with ``wide Jacobians''.

\subsubsection{Vector-Jacobian products (VJPs)}
Similarly, if we consider 
a perturbation to $\vec u\in\mathbb{R}^m$ to the output of the $f$, we obtain
the \textsl{vector-Jacobian product} (VJP):
\begin{equation}
(\vec{x}, \vec{u}) \mapsto \partial f(\vec{x})^{\intercal}\,\vec u
\label{eq:vjp}
\end{equation}
Whereas JVPs are useful for computing the Jacobian matrix one column at a time, we can use VJPs
to compute the Jacobian one row at a time:
\begin{equation}
    \vec v^\intercal\,\partial f(\vec x)= 
    (1\,0\, \cdots \, 0)\,
    \begin{pmatrix}
\frac{\partial f_{1}}{\partial x_{1}} & \cdots & \frac{\partial f_{1}}{\partial x_{n}} \\
\vdots & \ddots & \vdots \\
\frac{\partial f_{m}}{\partial x_{1}} & \cdots & \frac{\partial f_{m}}{\partial x_{n}}
    \end{pmatrix}
        = \begin{pmatrix}
        \frac{\partial f_{1}}{\partial x_{1}}\\
        \frac{\partial f_{1}}{\partial x_{2}}\\
        \vdots\\
        \frac{\partial f_{1}}{\partial x_{n}}
    \end{pmatrix}
\end{equation}
For a scalar function $f\,:\,\mathbb{R}^n\rightarrow\mathbb{R}$, the VJP is simply the gradient 
$\nabla f$:
\begin{equation}
    \vec v^\intercal\,\partial f(\vec x)\equiv \nabla f(\vec x)=
    \left(1\right)\,\left(  \frac{\partial f}{\partial x_{1}} \,\cdots\,\frac{\partial f}{\partial x_{n}}   \right)
    =\begin{pmatrix}
        \frac{\partial f}{\partial x_{1}}\\
        \vdots\\
        \frac{\partial f}{\partial x_{n}}
    \end{pmatrix}
\end{equation}

\subsubsection{Chain rule}
Let's now consider a function $f\,:\, \mathbb{R}^n\rightarrow\mathbb{R}^m$ which is a composition
three functions $a$, $b$ and $c$:
\begin{equation}
f=c \circ b \circ a
\end{equation}
The \textsl{chain rule} from calculus says that we can differentiate $f$ if we know how to differentiate 
each of its constituent parts:
\begin{align}
\partial f(\vec{x})&=\partial c(b(a(\vec{x}))) \circ \partial b(a(\vec{x})) \circ \partial a(\vec{x})\\
&=\partial c(b(a(\vec{x})))[\partial b(a(\vec{x}))[\partial a(\vec{x})]]
\label{eq:chain_rule}
\end{align}
The idea behind AD is to decompose complicated functions $f$ into constituent parts (sometimes 
called \textsl{primitive functions}) which we know how to differentiate  and then use the chain 
rule to obtain the derivative of $f$.
Equation~\ref{eq:chain_rule} consists of a series of matrix multiplications. Although the order
in which we multiply those matrices does not matter because of the associativity of matrix 
multiplication, it matters a great deal when we are implementing AD on a computer.
Let's say that we want to compute the JVP $\partial f(\vec{x})\,\vec v$ of this composite 
function:
\begin{equation}
 \partial f(\vec{x})\,\vec v = \partial c(b(a(\vec{x})))[\partial b(a(\vec{x}))[\partial a(\vec{x})\vec v]]
 \label{eq:jvp_chain_rule}
\end{equation}
Right-to-left evaluation of the expression on the right-hand side of 
Equation~\ref{eq:jvp_chain_rule} is known  as \textsl{forward-mode automatic differentiation}.
In forward-mode AD we evaluate \textsl{primal terms} $\vec{x}, a(\vec{x}), b(a(\vec{x}))$ 
alongside the \textsl{tangent terms} 
$\partial a(\vec{x}), \partial b(a(\vec{x})), \partial c(b(a(\vec{x})))$.

If instead we are interested in evaluating the VJP the order of operations is reversed:
\begin{equation}
\partial f(\vec{x})^{\intercal}\,\vec u=\partial a(\vec{x})^{\intercal}\left[\partial b(a(\vec{x}))^\intercal\left[\partial c(b(a(\vec{x})))^{\intercal}\vec u\right]\right]
\label{eq:vjp_chain_rule}
\end{equation}
To evaluate  Equation~\label{eq:vjp_chain_rule} we first have to traverse the whole chain 
computing the primals $\vec{x}, a(\vec{x}), b(a(\vec{x}))$ as we go and then we proceed in reverse order 
computing the tangents 
$\partial c(b(a(\vec{x})))^{\intercal}, \partial b(a(\vec{x}))^{\intercal}, \partial a(\vec{x})^{\intercal}$.
This process is known as \textsl{reverse-mode automatic differentiation} or 
\textsl{backpropagation} (since we are propagating the gradients backwards from the function output).

Using JVPs and VJPs is much more memory efficient than computing the Jacobian matrix directly 
because all intermediate results computed when applying the chain rule are 1D vectors that 
don't take up a lot of memory. The key difference between reverse mode and  forward mode is that 
in the case of forward-mode AD we can just evaluate the JVPs as we evaluate the function itself 
whereas in reverse-mode AD we need to store the primal terms for the entire chain before 
backpropagating so reverse-mode AD is more memory intensive.
Reverse-mode is usually far more popular than forward-mode in optimization and machine learning 
because the function $f$ is some cost function with  a high-dimensional input and a scalar 
output so we only need one VJP to obtain the entire Jacobian matrix. There is also mixed-mode 
automatic differentiation which combines forward-mode and reverse-mode to optimize memory use.

In the general case functions can be more complex than a simple chain of compositions and in AD 
frameworks they are represented as \textsl{Directed Acyclic Graphs} (DAGs), sometimes also 
called \textsl{computation graphs} or \textsl{circuits}. We point the reader to the Section 
6.2 in \citep{murphy_book_2023} for details on how to generalize AD to graphs and also higher 
order AD which is also possible using the same formalism.

\subsection{AD libraries}
The big challenge in automatic differentiation is creating a library wich can knows how to compute 
JVPs and VJPs for each primitive operation (for example, multiplication, division, 
matrix multiplication, array slicing, etc.).
Because the primitive operations need to be coded 
manually, expressing a given computation (say an astrophysics code written in \textsf{C++}) in 
an AD library often requires some effort and it is a bit like programming in a different 
language. 
Popular AD libraries are generally developed by large groups of researchers and with the backing 
of major tech companies.

There are two main kinds of AD libraries:
\begin{itemize}
    \item \textbf{Static graph libraries} 
    \item \textbf{Dynamic graph libraries}
\end{itemize}
The static graph libraries first build the entire computation graph for a given program before 
the graph is evaluated. Examples include \textsf{Theano}\footnote{Theano is no longer 
maintained but it has recently been ressurected and turned into \textsf{aesara} 
(\url{https://github.com/aesara-devs/aesara})} \citep{arXiv:1605.02688}, 
\textsf{TensorFlow}\footnote{\url{https://www.tensorflow.org/}} \citep{arXiv:1603.04467}, 
and the \textsf{Stan Math Library} \citep{arXiv:1509.07164}.
Dynamic graph libraries build the graph as the program is is being executed. Examples are 
\textsf{Tensorflow 2}, \textsf{JAX}\footnote{\url{https://jax.readthedocs.io/en/latest/}.} 
\citep{jax2018github} and \textsf{PyTorch}\footnote{\url{https://pytorch.org/}.} \citep{pytorch}.
These libraries have many features in addition to autodiff, most notably support for
for GPUs. They are written in Python but the primitive operations are usually 
expressed in a lower level language like \textsf{C++} and \textsf{CUDA C++} 
for GPU support.

The most recent and most research oriented of these libraries is Google's \textsf{JAX} 
library.
\textsf{JAX} is written in \textsf{Python} and it offers nearly identical functionallity 
to the widely popular \textsf{NumPy} \textsf{Python} library for manipulating 
multi-dimensional arrays. \textsf{JAX}takes Python code as an input, converts it 
to an intermediate language (IR) representation called \textsf{jaxpr} which is 
used to perform autodiff, and finally the resulting code for evaluating functions 
and their gradients is \textsl{Just-In-Time} (JIT) compiled to a lower level 
assembly like language called \textsf{HLO-IR} using the 
\textsf{XLA}\footnote{\url{https://www.tensorflow.org/xla}.} compiler in such a way 
that the code can be executed on different hardware architecutres such  as 
CPUs, GPUs and TPUs (Tensor Processing Units).
In practice this means that one can write \textsf{Python} code using a \textsf{NumPy} 
like library, easily compute gradients of the code through autodiff and get 
the performance of a compiled language like $C++$ through on-the-fly JIT 
compilation. \textsf{JAX} natively supports \textsf{Python} language constructs such 
as loops, conditionals and data structures such as tuples and dictionaries.
Almost everything you could write in \textsf{Python} is differentiable with 
\textsf{JAX} although there is no guarantee that the gradients are meaningful 
if the function represented by the code is not itself differentiable.
Implementing a real-world model in an AD framework often requires making individual 
functions ``smoother'' so they are more easily differentiable (see 
Chapter~\ref{ch:microlensing2} for more on this point).

Many of the statistical algorithms I have mentioned  in 
Section~\ref{ssec:sampling_methods} are now implemented in \textsf{JAX} and 
\textsf{JAX} has also been used within astronomy. \citet{2022ApJS..258...31K} built 
an autodifferentiable model for exoplanet spectrum modeling, 
the \textsf{exoplanet} code \citet{2021JOSS....6.3285F} for modeling transit light 
curves and radial velocity time series uses \textsf{JAX} as a backend, and 
\citet{2022arXiv220207663G} used \textsf{JAX} to build a differentiable model 
for strong lensing  which produces samples from a Bayesian posterior by  first
exploring the posterior using an optimizer, running variational inference to 
obtain an estimate of the posterior covariance followed by Hamiltonian Monte Carlo
to obtain the final samples. Importantly, \textsf{JAX} makes it relatively 
straightforward to implement custom primitives. One of the contributions of this 
thesis is an implementation of a custom JAX primitive for computing the roots 
of complex polynomials (see Chapter~\ref{ch:microlensing2} for more details).
The process of building custom primitives involves specifying the JVP rule 
(Equation~\ref{eq:jvp}). Interestingly, to support reverse-mode AD we it is 
not necessary to also specify the VJP rule (Equation~\ref{eq:vjp}) because 
\textsf{JAX} is able to deduce the transpose of the JVP automatically\footnote{
    Essentially, this is possible because the implementation of the JVP rule 
    in \textsf{Python} consists of several \textsf{JAX} primitives such as addition
    and multiplication and \textsf{JAX} already knows how to transpose those 
    primitives.} (see \citet{arXiv:2105.09469} for details).

\subsection{Probability + programming + autodiff = probabilistic programming}
The final topic in this chapter nicely ties up everything we have discussed so far.
The term \textsl{probabilistic programming} refers to the use of software libraries 
which  provide a high-level API for constructing statistical models (usually in 
the Bayesian paradigm). Specifically, 
the key components of modern probabilistic programming languages (PPLs) include a backend
autodiff framework, methods for initializing \emph{random variables} with 
associated pdf or pmf distribution functions and evaluation the (log) probability, 
and finally an one or several optimizers and samplers. The idea behind these 
frameworks is to provide an easy to use interface for constructing probabilistic 
models and to abstract away the inference as much as possible.
Why is this useful for scientists? Because these frameworks simplify the process  
of building probabilistic  models and because they include well-tested and high
performance implementations of state-of-the-art inference methods such as 
Hamiltonian Monte Carlo.

The most popular such library is \textsf{Stan} \citep{2017JSS....76....1C}. 
Stan is built in \textsf{C++}, it uses the \textsf{Stan Math Library} for automatic 
differentiation, it uses a variant of the NUTS HMC sampler and also has some 
support for Variational Inference. \textsf{Stan} is oriented primarily towards 
applications in the social sciences and it is not as flexible and easily extensible 
as some other PPLs. \textsf{PyMC} is another PPL which is very similar to 
\textsf{Stan} except that it is implemented in \textsf{Python} and it uses 
the \textsf{aesara}\footnote{\url{https://github.com/aesara-devs/aesara}.} library 
for AD (it used to be called \textsf{PyMC3} and use \textsf{Theano} for AD) and it 
also supports \textsf{JAX}. 
\textsf{PyMC} is somewhat more popular in the physical sciences and easier to 
extend  with custom functionality. The \textsf{exoplanet} code 
\citep{2021JOSS....6.3285F} is a good example of an astrophysics code which 
interfaces with \textsf{PyMC}.  There is also 
\textsf{numpyro}\footnote{\url{https://num.pyro.ai/}.}
\citep{arXiv:1912.11554} which is also built in \textsf{Python} and uses 
\textsf{JAX} for AD and \textsf{Tensorflow Probability} \citep{arXiv:1711.10604} 
which supports bot \textsf{TensorFlow} and \textsf{JAX} for AD. \textsf{PyMC}, \textsf{numpyro}
and \textsf{Tensorflow Probability} all use a variant of NUTS similar to the one 
implemented in \textsf{Stan} as the core algorithm for inference but also 
include support for Variational Inference and additional samplers. 
I use PPLs (mostly \textsf{numpyro}) extensively throughout the rest of this thesis.

% CHAPTER 3: Modeling single lens microlensing events 
%% PLOT POINTWISE LOO 
\chapter{Modeling single lens microlensing events}
\label{ch:microlensing}
% single lens events are important because...
\section{A differentiable single lens extended source model}
\section{Practical solutions for sampling multi-modal posteriors}
\section{Hierarchical inference}

% CHAPTER 4: Modeling binary and triple microlensing events
\chapter{Modeling binary and triple lens microlensing events}
\label{ch:microlensing2}
\section{A differentiable complex polynomial root solver}
\section{\textsf{caustics} -- computing the magnification of an extended limb-darkened source}
\section{Comparison with previous work}
\section{Modeling binary lens events}
\section{Population-level inference using hierarchical modeling}

% CHAPTER 5: Mapping the surface of Io 
\chapter{Mapping the surface of Io}
\label{ch:mapping_io}
\section{The data}
\section{Information content}
\section{Pixel sampling}
\section{A static map model}
\section{A dynamic map model}

% CHAPTER 6: Mapping the surfaces of exoplanets
\chapter{Mapping the surfaces of exoplanets}
\label{ch:mapping_exoplanets}
\section{Previous work}
\section{What can we learn using eclipse mapping?}
\section{Time dependent maps}

% CHAPTER 7: Conclusion
\chapter{Conclusion}
\section{Summary of the thesis}
\subsection{Microlensing}
\subsection{Occultation mapping}
\section{Future work}
\subsection{Microlensing in the era of the Roman telescope}
\subsection{Mapping volcanic activity on Io}
\subsection{Mapping exoplanets with JWST and beyond}
